---
title: "SC_source_guide"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Decedent Data

UCLA obtains custodial decedent data for South Carolina state prisons from public records requests to the South Carolina Department of Corrections.

Example files included in corresponding documents folder.

Some processing of the files takes place, including the selection of corresponding variables and linking data to UCLA datasets.  

```{r}
# Load SC decedent data
sc.deaths <- read_ucla_deaths(agencies = 'SC')

# Show variables present in SC decedent data
colnames(sc.deaths)

# UCLA.ID is present - join UCLA facility data
sc.facility <- sc.deaths %>%
                pull_ucla_fac_data()
head(sc.facility)

# Compare UCLA data with BJS data
compare_ucla_bjs() %>%
    subset(State == 'South Carolina') %>%
    arrange(desc(Year))

# Plot SC decedent data (wrangled to monthly numbers)
ggplot() +
    geom_line(data = sc.deaths %>% group_by(Year, Month) %>% summarise(Total.Deaths = n()) %>% 
                  mutate(Date = as.Date(str_c(Year,Month,'01'), format = '%Y%B%d')),
              aes(x = Date, y = Total.Deaths)
              )

```


## Demographic Data

UCLA currently does not have demographic data processed for South Carolina state prisons. Please wait for the next update to this data. 
