---
title: "IL_source_guide"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Environment

```{r}
# Load relevent packages and functions
library(tidyverse)
source('Code/general_utilities.R') # Note you may have to change your Rmarkdown option for evaluate chunks to project instead of document
options(scipen=10)
```

## Decedent Data

UCLA obtains custodial decedent data for Illinois state prisons from public records requests to the Illinois Department of Corrections.

Example files included in corresponding documents folder.

Some processing of the files takes place, including the selection of corresponding variables and linking data 
to UCLA datasets. 

```{r}
# Load IL decedent data
il.deaths <- read_CMP_deaths(agencies = 'IL')

# Show variables present in IL decedent data
colnames(il.deaths)

# UCLA.ID is present - join UCLA facility data
il.deaths %>%
    pull_CMP_fac_data()

# Compare CMP data with BJS data
compare_CMP_bjs(source = 'MCI') %>%
    subset(State == 'Illinois') %>%
    arrange(desc(Year))

# Plot IL decedent data 
ggplot() +
    geom_line(data = il.deaths %>% group_by(Year) %>% summarise(Total.Deaths = n()),aes(x = Year, y = Total.Deaths))

```


## Demographic Data

UCLA obtains custodial demographic data for Illinois state prisons from population datasets posted by the Illinois Department of Corrections (IDOC). Because IDOC provides population data on an individual level, different age and sex profiles can be constructed at will for dates of available data. One reconstruction, produced by our data team, is available here.

Link to report page: https://www2.illinois.gov/idoc/reportsandstatistics/Pages/Prison-Population-Data-Sets.aspx

Some processing of the files takes place. 

```{r}
# Load IL demographic data
il.demographics <- harmonize_CMP_dem(agencies = 'IL')

# Show IL demographic data to assess origin type
head(il.demographics)

# The origin type is combined. This means that age and sex groups were aggregated and reported together.  

# Plot IL demographic data 
ggplot() +
    geom_bar(data = il.demographics %>% group_by(Date, Age) %>% summarise(Number = sum(Number, na.rm = TRUE)),
             aes(x = Date, y = Number, fill = Age), stat = 'identity') +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) # To see dates on x axis



```

