---
title: "VT_source_guide"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Environment

```{r}
# Load relevent packages and functions
library(tidyverse)
source('Code/general_utilities.R') # Note you may have to change your Rmarkdown option for evaluate chunks to project instead of document
options(scipen=10)
```

## Decedent Data

UCLA obtains custodial decedent data for Vermont state prisons from public records requests to the Vermont Department of Corrections.

Example files included in corresponding documents folder. 

Some processing of the files takes place, including the selection of corresponding variables and linking data to UCLA datasets. 

```{r}
# Load VT decedent data
vt.deaths <- read_ucla_deaths(agencies = 'VT')

# Show variables present in VT decedent data
colnames(vt.deaths)

# UCLA.ID is not present - please wait for data update to join UCLA facility data

# Compare UCLA data with BJS data
compare_ucla_bjs() %>%
    subset(State == 'Vermont') %>%
    arrange(desc(Year))

# Plot VT decedent data (wrangled to monthly numbers)
ggplot() +
    geom_line(data = vt.deaths %>% group_by(Year, Month) %>% summarise(Total.Deaths = sum(Total.Deaths)) %>% 
                  mutate(Date = as.Date(str_c(Year,Month,'01'), format = '%Y%B%d')),
              aes(x = Date, y = Total.Deaths)
              ) +
    geom_smooth(data = vt.deaths %>% group_by(Year, Month) %>% summarise(Total.Deaths = sum(Total.Deaths)) %>% 
                  mutate(Date = as.Date(str_c(Year,Month,'01'), format = '%Y%B%d')),
              aes(x = Date, y = Total.Deaths)
              )

```


## Demographic Data

UCLA currently does not have demographic data processed for Vermont state prisons. Please wait for the next update to this data. 
