---
title: "OK_source_guide"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Environment

```{r}
# Load relevent packages and functions
library(tidyverse)
source('Code/general_utilities.R') # Note you may have to change your Rmarkdown option for evaluate chunks to project instead of document
options(scipen=10)
```

## Decedent Data

UCLA obtains custodial decedent data for Oklahoma state prisons from public records requests to the Oklahoma Department of Corrections.

Example files included in corresponding documents folder.

Some processing of the files takes place, including the selection of corresponding variables and linking data to UCLA datasets. 

```{r}
# Load OK decedent data
ok.deaths <- read_CMP_deaths(agencies = 'OK')

# Show variables present in OK decedent data
colnames(ok.deaths)

# UCLA.ID is present - join UCLA facility data
ok.facility <- ok.deaths %>%
               pull_CMP_fac_data()
head(ok.facility)

# Compare CMP data with BJS data
compare_CMP_bjs(source = 'MCI') %>%
    subset(State == 'Oklahoma') %>%
    arrange(desc(Year))

# Plot OK decedent data 
ggplot() +
    geom_line(data = ok.deaths %>% group_by(Year) %>% summarise(Total.Deaths = n()),aes(x = Year, y = Total.Deaths))

```


## Demographic Data

UCLA currently does not have demographic data processed for Oklahoma state prisons. Please wait for the next update to this data. 
