---
title: "WI_source_guide"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Environment

```{r}
# Load relevent packages and functions
library(tidyverse)
source('Code/general_utilities.R') # Note you may have to change your Rmarkdown option for evaluate chunks to project instead of document
options(scipen=10)
```

## Decedent Data

UCLA obtains custodial decedent data for Wisconsin state prisons from public records requests to the Wisconsin Department of Correction.

Example files included in corresponding documents folder.

Some processing of the files takes place, including the selection of corresponding variables and linking data to UCLA datasets. 

```{r}
# Load WI decedent data
wi.deaths <- read_CMP_deaths(agencies = 'WI')

# Show variables present in WI decedent data
colnames(wi.deaths)

# UCLA.ID is not present - please wait for data update to join UCLA facility data

# Compare CMP data with BJS data
compare_CMP_bjs(source = 'MCI') %>%
    subset(State == 'Wisconsin') %>%
    arrange(desc(Year))

# Plot WI decedent data 
ggplot() +
    geom_line(data = wi.deaths %>% group_by(Year) %>% summarise(Total.Deaths = n()),aes(x = Year, y = Total.Deaths))

```


## Demographic Data

UCLA obtains custodial demographic data for Wisconsin state prisons from annual population reports from the Wisconsin Department of Correction.

Example files in corresponding folder and links to reports in raw data file.

Some processing of the files takes place. 

```{r}
# Load WI demographic data
wi.demographics <- harmonize_CMP_dem(agencies = 'WI')

# Show WI demographic data to assess origin type
head(wi.demographics)

# The origin type is distinct. This means that age groups were reported seperately from sex groups for demographic data. In order to construct age and sex groups for a date from this distinct data, we applied reported sex ratios to each age group. This assumption is likely incorrect.

# Plot MA demographic data 
ggplot() +
    geom_bar(data = wi.demographics %>% group_by(Date, Age) %>% summarise(Number = sum(Number, na.rm = TRUE)),
             aes(x = Date, y = Number, fill = Age), stat = 'identity') +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) # To see dates on x axis



```
