---
title: "KS_source_guide"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Environment

```{r}
# Load relevent packages and functions
library(tidyverse)
source('Code/general_utilities.R') # Note you may have to change your Rmarkdown option for evaluate chunks to project instead of document
options(scipen=10)
```

## Decedent Data

UCLA obtains custodial decedent data for Kansas state prisons from public records requests to the Kansas Department of Corrections.

Example files included in corresponding documents folder. Partial data for 2020 is present in example files. Data is currently only cleaned for 2018-2019.

Some processing of the files takes place, including the selection of corresponding variables and linking data 
to UCLA datasets. 

```{r}
# Load KS decedent data
ks.deaths <- read_ucla_deaths(agencies = 'KS')

# Show variables present in NM decedent data
colnames(ks.deaths)

# UCLA.ID is present - join UCLA facility data
ks.deaths %>%
    pull_ucla_fac_data()

# Compare UCLA data with BJS data
compare_ucla_bjs() %>%
    subset(State == 'Kansas') %>%
    arrange(desc(Year))

# Plot KS decedent data (wrangled to monthly numbers)
ggplot() +
    geom_line(data = ks.deaths %>% group_by(Year, Month) %>% summarise(Total.Deaths = n()) %>% 
                  mutate(Date = as.Date(str_c(Year,Month,'01'), format = '%Y%B%d')),
              aes(x = Date, y = Total.Deaths)
              )

```


## Demographic Data

UCLA currently does not have demographic data processed for Kansas state prisons. Please wait for the next update to this data. 


