---
title: "NC_source_guide"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Environment

```{r}
# Load relevent packages and functions
library(tidyverse)
source('Code/general_utilities.R') # Note you may have to change your Rmarkdown option for evaluate chunks to project instead of document
options(scipen=10)
```

## Decedent Data

UCLA obtains custodial decedent data for North Carolina state prisons from public records requests to the North Carolina Department of Public Safety.

Example files included in corresponding documents folder.

Some processing of the files takes place, including the selection of corresponding variables and linking data to UCLA datasets. 

```{r}
# Load NC decedent data
nc.deaths <- harmonize_ucla_deaths(agencies = 'NC')

# Show variables present in NC decedent data
colnames(nc.deaths)

# UCLA.ID is present - join UCLA facility data
nc.deaths %>%
    pull_ucla_fac_data()

# Compare UCLA data with BJS data
compare_ucla_bjs() %>%
    subset(State == 'North Carolina') %>%
    arrange(desc(Year))

# Plot NC decedent data (wrangled to monthly numbers)
ggplot() +
    geom_line(data = nc.deaths %>% group_by(Year, Month) %>% summarise(Total.Deaths = n()) %>% 
                  mutate(Date = as.Date(str_c(Year,Month,'01'), format = '%Y%B%d')),
              aes(x = Date, y = Total.Deaths)
              ) +
    geom_smooth(data = nc.deaths %>% group_by(Year, Month) %>% summarise(Total.Deaths = n()) %>% 
                  mutate(Date = as.Date(str_c(Year,Month,'01'), format = '%Y%B%d')),
              aes(x = Date, y = Total.Deaths)
              )

```


## Demographic Data

UCLA obtains custodial demographic data for North Carolina state prisons from population datasets posted by the North Carolina Department of Public Safety (NCDPS). Because NCDPS provides population data on an individual level, different age and sex profiles can be constructed at will for dates of available data. One reconstruction, produced by our data team, is available here.


Some processing of the files takes place. 

```{r}
# Load NC demographic data
nc.demographics <- harmonize_ucla_dem(agencies = 'NC')

# Show NC demographic data to assess origin type
head(nc.demographics)

# The origin type is combined. This means that age and sex groups were aggregated and reported together.  

# Plot NC demographic data 
ggplot() +
    geom_bar(data = nc.demographics %>% group_by(Date, Age) %>% summarise(Number = sum(Number, na.rm = TRUE)),
             aes(x = Date, y = Number, fill = Age), stat = 'identity') +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) # To see dates on x axis



```