---
title: "PA_source_guide"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Environment

```{r}
# Load relevent packages and functions
library(tidyverse)
source('Code/general_utilities.R') # Note you may have to change your Rmarkdown option for evaluate chunks to project instead of document
options(scipen=10)
```

## Decedent Data

UCLA obtains custodial decedent data for Pennsylvania state prisons from public records requests to the Pennsylvania Department of Corrections.

Example files included in corresponding documents folder.

Some processing of the files takes place, including the selection of corresponding variables and linking data 
to UCLA datasets. 

```{r}
# Load PA decedent data
pa.deaths <- harmonize_CMP_deaths(agencies = 'PA')

# Show variables present in PA decedent data
colnames(pa.deaths)

# UCLA.ID is present - join UCLA facility data
pa.deaths %>%
    pull_CMP_fac_data()

# Compare CMP data with BJS data
compare_CMP_bjs(source = 'MCI') %>%
    subset(State == 'Pennsylvania') %>%
    arrange(desc(Year))

# Plot PA decedent data 
ggplot() +
    geom_line(data = pa.deaths %>% group_by(Year) %>% summarise(Total.Deaths = n()),aes(x = Year, y = Total.Deaths))

```


## Demographic Data

UCLA obtains custodial demographic data for Pennsylvania state prisons from monthly population reports from the Pennsylvania Department of Corrections.

Example files in corresponding folder and links to reports in raw data file.

Some processing of the files takes place. 

```{r}
# Load PA demographic data
pa.demographics <- harmonize_CMP_dem(agencies = 'PA')

# Show PA demographic data to assess origin type
head(pa.demographics)

# The origin type is distinct. This means that age groups were reported seperately from sex groups for demographic data. In order to construct age and sex groups for a date from this distinct data, we applied reported sex ratios to each age group. This assumption is likely incorrect.  

# Plot PA demographic data 
ggplot() +
    geom_bar(data = pa.demographics %>% group_by(Date, Age) %>% summarise(Number = sum(Number, na.rm = TRUE)),
             aes(x = Date, y = Number, fill = Age), stat = 'identity') +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) # To see dates on x axis



```