---
title: "CO_source_guide"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Environment

```{r}
# Load relevent packages and functions
library(tidyverse)
source('Code/general_utilities.R') # Note you may have to change your Rmarkdown option for evaluate chunks to project instead of document
options(scipen=10)
```

## Decedent Data

UCLA obtains custodial decedent data for Colorado state prisons from public records requests to the Colorado Department of Corrections.

Example files included in corresponding documents folder.

Some processing of the files takes place, including the selection of corresponding variables and linking data 
to UCLA datasets. 

```{r}
# Load CO decedent data
co.deaths <- harmonize_ucla_deaths(agencies = 'CO')

# Show variables present in CO decedent data
colnames(co.deaths)

# UCLA.ID is present - join UCLA facility data
co.facility <- co.deaths %>%
                pull_ucla_fac_data()
head(co.facility)

# Compare UCLA data with BJS data
compare_ucla_bjs() %>%
    subset(State == 'Colorado') %>%
    arrange(desc(Year))

# Plot CO decedent data (wrangled to monthly numbers)
ggplot() +
    geom_line(data = co.deaths %>% group_by(Year, Month) %>% summarise(Total.Deaths = n()) %>% 
                  mutate(Date = as.Date(str_c(Year,Month,'01'), format = '%Y%B%d')),
              aes(x = Date, y = Total.Deaths)
              ) +
    geom_smooth(data = co.deaths %>% group_by(Year, Month) %>% summarise(Total.Deaths = n()) %>% 
                  mutate(Date = as.Date(str_c(Year,Month,'01'), format = '%Y%B%d')),
              aes(x = Date, y = Total.Deaths)
              )

```


## Demographic Data

UCLA obtains custodial demographic data for Colorado state prisons from annual population appendices from the Colorado Department of Corrections.

Example files in corresponding folder and links to reports in raw data file.

Some processing of the files takes place. 

```{r}
# Load CO demographic data
co.demographics <- harmonize_ucla_dem(agencies = 'CO')

# Show CO demographic data to assess origin type
head(co.demographics)

# The origin type is combined. This means that age and sex groups were aggregated and reported together.   

# Plot CO demographic data 
ggplot() +
    geom_bar(data = co.demographics %>% group_by(Date, Age) %>% summarise(Number = sum(Number, na.rm = TRUE)),
             aes(x = Date, y = Number, fill = Age), stat = 'identity') +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) # To see dates on x axis



```
