---
title: "MN_source_guide"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Decedent Data

UCLA obtains custodial decedent data for Minnesota state prisons from public records requests to the Minnesota Department of Corrections.

Example files included in corresponding documents folder.

Some processing of the files takes place, including the selection of corresponding variables and linking data to UCLA datasets. 

```{r}
# Load MN decedent data
mn.deaths <- read_CMP_deaths(agencies = 'MN')

# Show variables present in MN decedent data
colnames(mn.deaths)

# UCLA.ID is present - join UCLA facility data
mn.facility <- mn.deaths %>%
                pull_ucla_fac_data()
head(mn.facility)

# Compare CMP data with BJS data
compare_CMP_bjs(source = 'MCI') %>%
    subset(State == 'Minnesota') %>%
    arrange(desc(Year))

# Plot MN decedent data 
ggplot() +
    geom_line(data = mn.deaths %>% group_by(Year) %>% summarise(Total.Deaths = n()),aes(x = Year, y = Total.Deaths))

```


## Demographic Data

UCLA currently does not have demographic data processed for Minnesota state prisons. Please wait for the next update to this data. 
