---
title: "MD_source_guide"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Decedent Data

UCLA obtains custodial decedent data for Maryland state prisons from public records requests to the Maryland Department of Corrections.

Example files included in corresponding documents folder.

Some processing of the files takes place, including the selection of corresponding variables and linking data to UCLA datasets. 

```{r}
# Load MD decedent data
md.deaths <- harmonize_CMP_deaths(agencies = 'MD')

# Show variables present in MD decedent data
colnames(md.deaths)

# UCLA.ID is present - join UCLA facility data
md.facility <- md.deaths %>%
                pull_CMP_fac_data()
head(md.facility)

# Compare CMP data with BJS data
compare_CMP_bjs(source = 'MCI') %>%
    subset(State == 'Maryland') %>%
    arrange(desc(Year))

# Plot MD decedent data 
ggplot() +
    geom_line(data = md.deaths %>% group_by(Year) %>% summarise(Total.Deaths = n()),aes(x = Year, y = Total.Deaths))

```


## Demographic Data

UCLA obtains custodial demographic data for Maryland state prisons from annual population reports from the Maryland Department of Corrections.

Example files in corresponding folder and links to reports in raw data file.

Some processing of the files takes place. 

```{r}
# Load MD demographic data
md.demographics <- harmonize_CMP_dem(agencies = 'MD')

# Show MD demographic data to assess origin type
head(md.demographics)

# The origin type is distinct. This means that age groups were reported seperately from sex groups for demographic data. In order to construct age and sex groups for a date from this distinct data, we applied reported sex ratios to each age group. This assumption is likely incorrect.   

# Plot MD demographic data 
ggplot() +
    geom_bar(data = md.demographics %>% group_by(Date, Age) %>% summarise(Number = sum(Number, na.rm = TRUE)),
             aes(x = Date, y = Number, fill = Age), stat = 'identity') +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) # To see dates on x axis



```
