---
title: "CA_source_guide"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Environment

```{r}
# Load relevent packages and functions
library(tidyverse)
source('Code/general_utilities.R') # Note you may have to change your Rmarkdown option for evaluate chunks to project instead of document
options(scipen=10)
```

## Decedent Data

UCLA obtains custodial decedent data for California state prisons from the California Department of Justice OpenJustice Data Portal.

Link to data portal: https://openjustice.doj.ca.gov/data

Some processing of the files takes place, including the selection of CDCR observations and corresponding variables and linking data to UCLA datasets. CA DOJ's deaths-in-custody data files were subset to 'CDC/CRC' observations in the 'custodial_responsibility_at_time_of_death' field. Data was last pulled from CA DOJ in July 2022 and data files were clipped to only contain deaths for full years (i.e. up to 2020).

```{r}
# Load CA decedent data
ca.deaths <- read_CMP_deaths(agencies = 'CA')

# Show variables present in CA decedent data
colnames(ca.deaths)

# UCLA.ID is present - join UCLA facility data
ca.deaths %>%
    pull_CMP_fac_data()

# Compare CMP data with BJS data
compare_CMP_bjs(source = 'MCI') %>%
    subset(State == 'California') %>%
    arrange(desc(Year))

# Plot CA decedent data 
ggplot() +
    geom_line(data = ca.deaths %>% group_by(Year) %>% summarise(Total.Deaths = n()),aes(x = Year, y = Total.Deaths))

```


## Demographic Data

UCLA obtains custodial demographic data for California state prisons from bi-annual population reports from the California Department of Corrections and Rehabilitation (CDCR).

Link to report page: https://www.cdcr.ca.gov/research/offender-outcomes-characteristics/offender-data-points/

Some processing of the files takes place. Example files included in corresponding documents folder. 

```{r}
# Load CA demographic data
ca.demographics <- harmonize_CMP_dem(agencies = 'CA')

# Show CA demographic data to assess origin type
head(ca.demographics)

# The origin type is distinct. This means that age groups were reported separately from sex groups for demographic data. In order to construct age and sex groups for a date from this distinct data, we applied reported sex ratios to each age group. This assumption is likely incorrect.

# Plot CA demographic data 
ggplot() +
    geom_bar(data = ca.demographics %>% group_by(Date, Age) %>% summarise(Number = sum(Number, na.rm = TRUE)),
             aes(x = Date, y = Number, fill = Age), stat = 'identity')+
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) # To see dates on x axis



```




