---
title: "ND_source_guide"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Decedent Data

UCLA obtains custodial decedent data for North Dakota state prisons from public records requests to the North Dakota Department of Corrections.

Example files included in corresponding documents folder.

Some processing of the files takes place, including the selection of corresponding variables and linking data to UCLA datasets. 

```{r}
# Load ND decedent data
nd.deaths <- read_ucla_deaths(agencies = 'ND')

# Show variables present in ND decedent data
colnames(nd.deaths)

# No UCLA.ID is present 

# Compare UCLA data with BJS data
compare_ucla_bjs() %>%
    subset(State == 'North Dakota') %>%
    arrange(desc(Year))

# Plot ND decedent data (wrangled to monthly numbers)
ggplot() +
    geom_line(data = nd.deaths %>% group_by(Year, Month) %>% summarise(Total.Deaths = n()) %>% 
                  mutate(Date = as.Date(str_c(Year,Month,'01'), format = '%Y%B%d')),
              aes(x = Date, y = Total.Deaths)
              )

```


## Demographic Data

UCLA currently does not have demographic data processed for North Dakota state prisons. Please wait for the next update to this data. 
