---
title: "TX_source_guide"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Environment

```{r}
# Load relevent packages and functions
library(tidyverse)
source('Code/general_utilities.R') # Note you may have to change your Rmarkdown option for evaluate chunks to project instead of document
options(scipen=10)
```

## Decedent Data

UCLA obtains custodial decedent data for Texas state prisons from the Texas Justice Intitiative. All credit for data production and maintainence goes to TJI and any use of the Texas decedent data contained in this dataset should follow their and our data license agreements.

Link to TJI custodial death dataset: https://texasjusticeinitiative.org/datasets/custodial-deaths

Some processing of the files takes place, including the selection of corresponding variables and linking data to project datasets. TJI's deaths-in-custody data files were subset to 'TDCJ' observations in the 'specific_type_of_custody_facility' field. The project used address information provided by TJI to crosswalk facility information to deaths. The crosswalk containing which facilities were joined to address spellings is currently saved in this documentation folder as `tdcj_tji_address_crosswalk.csv.` Data was last pulled from TJI in July 2022 and data files were clipped to only contain deaths for full years (i.e. up to 2021).

```{r}
# Load TX decedent data
tx.deaths <- read_CMP_deaths(agencies = 'TX')

# Show variables present in TX decedent data
colnames(tx.deaths)

# UCLA.ID is present - join UCLA facility data
tx.deaths %>%
    pull_CMP_fac_data()

# Compare CMP data with BJS data
compare_CMP_bjs(source = 'MCI') %>%
    subset(State == 'Texas') %>%
    arrange(desc(Year))

# Plot TX decedent data 
ggplot() +
    geom_line(data = tx.deaths %>% group_by(Year) %>% summarise(Total.Deaths = n()),aes(x = Year, y = Total.Deaths))

```


## Demographic Data

UCLA obtains custodial demographic data for Texas state prisons from population datasets posted by the Texas Department of Criminal Justice (TDCJ). Because TDCJ provides population data on an individual level, different age and sex profiles can be constructed at will for dates of available data. One reconstruction, produced by our data team, is available here.

Link to report page: https://www.tdcj.texas.gov/kss_inside.html

Some processing of the files takes place. 

```{r}
# Load TX demographic data
tx.demographics <- harmonize_CMP_dem(agencies = 'TX')

# Show TX demographic data to assess origin type
head(tx.demographics)

# The origin type is combined. This means that age and sex groups were aggregated and reported together.  

# Plot TX demographic data 
ggplot() +
    geom_bar(data = tx.demographics %>% group_by(Date, Age) %>% summarise(Number = sum(Number, na.rm = TRUE)),
             aes(x = Date, y = Number, fill = Age), stat = 'identity') +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) # To see dates on x axis



```

