---
title: "AL_source_guide"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Environment

```{r}
# Load relevent packages and functions
library(tidyverse)
source('Code/general_utilities.R') # Note you may have to change your Rmarkdown option for evaluate chunks to project instead of document
options(scipen=10)
```

## Decedent Data

UCLA obtains custodial decedent data for Alabama state prisons from monthly summary reports posted online by the Alabama Department of Corrections.

Example files included in corresponding documents folder. 

Some processing of the files takes place, including the selection of corresponding variables and linking data to UCLA datasets. 

```{r}
# Load AL decedent data
al.deaths <- read_ucla_deaths(agencies = 'AL')

# Show variables present in AL decedent data
colnames(al.deaths)

# UCLA.ID is not present - please wait for data update to join UCLA facility data

# Compare UCLA data with BJS data
compare_ucla_bjs() %>%
    subset(State == 'Alabama') %>%
    arrange(desc(Year))

# Plot AL decedent data (wrangled to monthly numbers)
ggplot() +
    geom_line(data = al.deaths %>% group_by(Year, Month) %>% summarise(Total.Deaths = n()) %>% 
                  mutate(Date = as.Date(str_c(Year,Month,'01'), format = '%Y%B%d')),
              aes(x = Date, y = Total.Deaths)
              ) +
    geom_smooth(data = al.deaths %>% group_by(Year, Month) %>% summarise(Total.Deaths = n()) %>% 
                  mutate(Date = as.Date(str_c(Year,Month,'01'), format = '%Y%B%d')),
              aes(x = Date, y = Total.Deaths)
              )

```


## Demographic Data

UCLA currently does not have demographic data processed for Alabama state prisons. Please wait for the next update to this data. 

