---
title: "MS_source_guide"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Environment

```{r}
# Load relevent packages and functions
library(tidyverse)
source('Code/general_utilities.R') # Note you may have to change your Rmarkdown option for evaluate chunks to project instead of document
options(scipen=10)
```

## Decedent Data

UCLA obtains custodial decedent data for Mississippi state prisons from public records requests to the Mississippi Department of Corrections.

Example files included in corresponding documents folder. 

Some processing of the files takes place, including the selection of corresponding variables and linking data to UCLA datasets. 

```{r}
# Load MS decedent data
ms.deaths <- read_ucla_deaths(agencies = 'MS')

# Show variables present in MS decedent data
colnames(ms.deaths)

# UCLA.ID is present - join UCLA facility data
ms.deaths %>%
    pull_ucla_fac_data()

# Compare UCLA data with BJS data
compare_ucla_bjs() %>%
    subset(State == 'Mississippi') %>%
    arrange(desc(Year))

# Plot MS decedent data (wrangled to monthly numbers)
ggplot() +
    geom_line(data = ms.deaths %>% group_by(Year, Month) %>% summarise(Total.Deaths = n()) %>% 
                  mutate(Date = as.Date(str_c(Year,Month,'01'), format = '%Y%B%d')),
              aes(x = Date, y = Total.Deaths)
              ) +
    geom_smooth(data = ms.deaths %>% group_by(Year, Month) %>% summarise(Total.Deaths = n()) %>% 
                  mutate(Date = as.Date(str_c(Year,Month,'01'), format = '%Y%B%d')),
              aes(x = Date, y = Total.Deaths)
              )

```


## Demographic Data

UCLA currently does not have demographic data processed for Mississippi state prisons. Please wait for the next update to this data. 
