---
title: "AZ_source_guide"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Environment

```{r}
# Load relevent packages and functions
library(tidyverse)
source('Code/general_utilities.R') # Note you may have to change your Rmarkdown option for evaluate chunks to project instead of document
options(scipen=10)
```

## Decedent Data

UCLA obtains custodial decedent data for Arizona state prisons from death notifications posted by the Arizona Department of Corrections.

Example files included in corresponding documents folder.

Some processing of the files takes place, including the selection of corresponding variables and linking data to UCLA datasets. 

```{r}
# Load AZ decedent data
az.deaths <- read_CMP_deaths(agencies = 'AZ')

# Show variables present in AZ decedent data
colnames(az.deaths)

# UCLA.ID is present - join UCLA facility data
az.facility <- az.deaths %>%
                pull_ucla_fac_data()
head(az.facility)

# Compare CMP data with BJS data
compare_CMP_bjs(source = 'MCI') %>%
    subset(State == 'Arizona') %>%
    arrange(desc(Year))

# Plot AZ decedent data 
ggplot() +
    geom_line(data = az.deaths %>% group_by(Year) %>% summarise(Total.Deaths = n()),aes(x = Year, y = Total.Deaths))

```


## Demographic Data

UCLA obtains custodial demographic data for Arizona state prisons from annual population reports from the Arizona Department of Corrections.

Example files in corresponding folder and links to reports in raw data file.

Some processing of the files takes place. 

```{r}
# Load AZ demographic data
az.demographics <- harmonize_CMP_dem(agencies = 'AZ')

# Show AZ demographic data to assess origin type
head(az.demographics)

# The origin type is combined. This means that age and sex groups were aggregated and reported together.   

# Plot AZ demographic data 
ggplot() +
    geom_bar(data = az.demographics %>% group_by(Date, Age) %>% summarise(Number = sum(Number, na.rm = TRUE)),
             aes(x = Date, y = Number, fill = Age), stat = 'identity') +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) # To see dates on x axis



```
