---
title: "AK_source_guide"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Environment

```{r}
# Load relevant packages and functions
library(tidyverse)
source('Code/general_utilities.R') 
options(scipen=10)
```

## Decedent Data

UCLA obtains custodial decedent data for Alaska state prisons from public records requests to the Alaska Department of Corrections.

Example files included in corresponding documents folder. Partial data for 2020 is present in example files. Data is currently only cleaned for 2018-2019.

Some processing of the files takes place, including the selection of corresponding variables and linking data 
to UCLA datasets. 

```{r}
# Load AK decedent data
ak.deaths <- read_CMP_deaths(agencies = 'AK')

# Show variables present in AK decedent data
colnames(ak.deaths)

# UCLA.ID is not present - please wait for data update to join UCLA facility data

# Compare CMP data with BJS data
compare_CMP_bjs(source = 'MCI') %>%
    subset(State == 'Alaska') %>%
    arrange(desc(Year))

# Plot AK decedent data 
ggplot() +
    geom_line(data = ak.deaths %>% group_by(Year) %>% summarise(Total.Deaths = n()),aes(x = Year, y = Total.Deaths))
    

```


## Demographic Data

UCLA currently does not have demographic data processed for Alaska state prisons. Please wait for the next update to this data. 