---
title: "NJ_source_guide"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Environment

```{r}
# Load relevent packages and functions
library(tidyverse)
source('Code/general_utilities.R') # Note you may have to change your Rmarkdown option for evaluate chunks to project instead of document
options(scipen=10)
```

## Decedent Data

UCLA obtains custodial decedent data for New Jersey state prisons from public records requests to the New Jersey Department of Corrections.

Example files included in corresponding documents folder.

Some processing of the files takes place, including the selection of corresponding variables and linking data to UCLA datasets. 

```{r}
# Load NJ decedent data
nj.deaths <- read_CMP_deaths(agencies = 'NJ')

# Show variables present in NJ decedent data
colnames(nj.deaths)

# UCLA.ID is present - join UCLA facility data
nj.deaths %>%
    pull_CMP_fac_data()

# Compare CMP data with BJS data
compare_CMP_bjs(source = 'MCI') %>%
    subset(State == 'New Jersey') %>%
    arrange(desc(Year))

# Plot NJ decedent data 
ggplot() +
    geom_line(data = nj.deaths %>% group_by(Year) %>% summarise(Total.Deaths = n()),aes(x = Year, y = Total.Deaths))

```


## Demographic Data

UCLA obtains custodial demographic data for New Jersey state prisons from monthly population reports from the New Jersey Department of Corrections.

Example files in corresponding folder and links to reports in raw data file.

Some processing of the files takes place. 

```{r}
# Load NJ demographic data
nj.demographics <- harmonize_CMP_dem(agencies = 'NJ')

# Show NJ demographic data to assess origin type
head(nj.demographics)

# The origin type is combined. This means that age and sex groups were aggregated and reported together.  

# Plot NJ demographic data 
ggplot() +
    geom_bar(data = nj.demographics %>% group_by(Date, Age) %>% summarise(Number = sum(Number, na.rm = TRUE)),
             aes(x = Date, y = Number, fill = Age), stat = 'identity') +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) # To see dates on x axis



```
