---
title: "NY_source_guide"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Environment

```{r}
# Load relevent packages and functions
library(tidyverse)
source('Code/general_utilities.R') # Note you may have to change your Rmarkdown option for evaluate chunks to project instead of document
options(scipen=10)
```

## Decedent Data

UCLA obtains custodial decedent data for New York state prisons from public records requests to the New York Department of Corrections and Community Supervision.

Example files included in corresponding documents folder.

Some processing of the files takes place, including the selection of corresponding variables and linking data 
to UCLA datasets. 

```{r}
# Load NY decedent data
ny.deaths <- harmonize_ucla_deaths(agencies = 'NY')

# Show variables present in NY decedent data
colnames(ny.deaths)

# UCLA.ID is present - join UCLA facility data
ny.deaths %>%
    pull_ucla_fac_data()

# Compare UCLA data with BJS data
compare_ucla_bjs() %>%
    subset(State == 'New York') %>%
    arrange(desc(Year))

# Plot NY decedent data (wrangled to monthly numbers)
ggplot() +
    geom_line(data = ny.deaths %>% group_by(Year, Month) %>% summarise(Total.Deaths = n()) %>% 
                  mutate(Date = as.Date(str_c(Year,Month,'01'), format = '%Y%B%d')),
              aes(x = Date, y = Total.Deaths)
              )

```


## Demographic Data

UCLA obtains custodial demographic data for New York state prisons from monthly population reports from the New York Department of Corrections and Community Supervision.

Example files in corresponding folder and links to reports in raw data file.

Some processing of the files takes place. 

```{r}
# Load NY demographic data
ny.demographics <- harmonize_ucla_dem(agencies = 'NY')

# Show NY demographic data to assess origin type
head(ny.demographics)

# The origin type is combined. This means that age and sex groups were aggregated and reported together.  

# Plot NY demographic data 
ggplot() +
    geom_bar(data = ny.demographics %>% group_by(Date, Age) %>% summarise(Number = sum(Number, na.rm = TRUE)),
             aes(x = Date, y = Number, fill = Age), stat = 'identity') +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) # To see dates on x axis



```

