---
title: "FL_source_guide"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Environment

```{r}
# Load relevent packages and functions
library(tidyverse)
source('Code/general_utilities.R') # Note you may have to change your Rmarkdown option for evaluate chunks to project instead of document
options(scipen=10)
```

## Decedent Data

UCLA obtains custodial decedent data for Florida state prisons from public records posted by the Florida Department of Corrections (FLDOC) on its OBIS search database.

Link to FLDOC OBIS database: http://www.dc.state.fl.us/OffenderSearch/InmateInfoMenu.aspx

Some processing of the files takes place, including the selection of corresponding variables and linking data 
to UCLA datasets. 

```{r}
# Load FL decedent data
fl.deaths <- read_CMP_deaths(agencies = 'FL')

# Show variables present in FL decedent data
colnames(fl.deaths)

# UCLA.ID is present - join UCLA facility data
fl.deaths %>%
    pull_CMP_fac_data()

# Compare CMP data with BJS data
compare_CMP_bjs(source = 'MCI') %>%
    subset(State == 'Florida') %>%
    arrange(desc(Year))

# Plot FL decedent data 
ggplot() +
    geom_line(data = fl.deaths %>% group_by(Year) %>% summarise(Total.Deaths = n()),aes(x = Year, y = Total.Deaths))

```


## Demographic Data

UCLA obtains custodial demographic data for Florida state prisons from public records posted by the Florida Department of Corrections on its OBIS search database. Because FLDOC provides population data on an individual level, different age and sex profiles can be constructed at will for dates of available data. One reconstruction, produced by our data team, is available here.

Link to FLDOC OBIS database: http://www.dc.state.fl.us/OffenderSearch/InmateInfoMenu.aspx

Some processing of the files takes place. 

```{r}
# Load FL demographic data
fl.demographics <- harmonize_CMP_dem(agencies = 'FL')

# Show FL demographic data to assess origin type
head(fl.demographics)

# The origin type is combined. This means that age and sex groups were aggregated and reported together.  

# Plot FL demographic data 
ggplot() +
    geom_bar(data = fl.demographics %>% group_by(Date, Age) %>% summarise(Number = sum(Number, na.rm = TRUE)),
             aes(x = Date, y = Number, fill = Age), stat = 'identity') +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) # To see dates on x axis



```

