---
title: "UT_source_guide"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Decedent Data

UCLA obtains custodial decedent data for Utah state prisons from public records requests to the Utah Department of Corrections.

Example files included in corresponding documents folder.

Some processing of the files takes place, including the selection of corresponding variables and linking data to UCLA datasets. 

```{r}
# Load UT decedent data
ut.deaths <- read_ucla_deaths(agencies = 'UT')

# Show variables present in UT decedent data
colnames(ut.deaths)

# No UCLA.ID is present 

# Compare UCLA data with BJS data
compare_ucla_bjs() %>%
    subset(State == 'Utah') %>%
    arrange(desc(Year))

# Plot UT decedent data (wrangled to monthly numbers)
ggplot() +
    geom_line(data = ut.deaths %>% group_by(Year, Month) %>% summarise(Total.Deaths = n()) %>% 
                  mutate(Date = as.Date(str_c(Year,Month,'01'), format = '%Y%B%d')),
              aes(x = Date, y = Total.Deaths)
              )

```


## Demographic Data

UCLA currently does not have demographic data processed for Utah state prisons. Please wait for the next update to this data. 
