---
title: "OH_source_guide"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Environment

```{r}
# Load relevent packages and functions
library(tidyverse)
source('Code/general_utilities.R') # Note you may have to change your Rmarkdown option for evaluate chunks to project instead of document
options(scipen=10)
```

## Decedent Data

UCLA obtains custodial decedent data for Ohio state prisons from public records requests to the Ohio Department of Corrections.

Example files included in corresponding documents folder. 

Some processing of the files takes place, including the selection of corresponding variables and linking data to UCLA datasets. 

```{r}
# Load OH decedent data
oh.deaths <- read_ucla_deaths(agencies = 'OH')

# Show variables present in OH decedent data
colnames(oh.deaths)

# UCLA.ID is present - join UCLA facility data
oh.deaths %>%
    pull_ucla_fac_data()

# Compare UCLA data with BJS data
compare_ucla_bjs() %>%
    subset(State == 'Ohio') %>%
    arrange(desc(Year))

# Plot OH decedent data (wrangled to monthly numbers)
ggplot() +
    geom_line(data = oh.deaths %>% group_by(Year, Month) %>% summarise(Total.Deaths = n()) %>% 
                  mutate(Date = as.Date(str_c(Year,Month,'01'), format = '%Y%B%d')),
              aes(x = Date, y = Total.Deaths)
              ) +
    geom_smooth(data = oh.deaths %>% group_by(Year, Month) %>% summarise(Total.Deaths = n()) %>% 
                  mutate(Date = as.Date(str_c(Year,Month,'01'), format = '%Y%B%d')),
              aes(x = Date, y = Total.Deaths)
              )

```


## Demographic Data

UCLA currently does not have demographic data processed for Ohio state prisons. Please wait for the next update to this data. 
