---
title: "MA_source_guide"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Environment

```{r}
# Load relevent packages and functions
library(tidyverse)
source('Code/general_utilities.R') # Note you may have to change your Rmarkdown option for evaluate chunks to project instead of document
options(scipen=10)
```

## Decedent Data

UCLA obtains custodial decedent data for Massachusetts state prisons from public records requests to the Massachusetts Department of Correction.

Example files included in corresponding documents folder.

Some processing of the files takes place, including the selection of corresponding variables and linking data to UCLA datasets. 

```{r}
# Load MA decedent data
ma.deaths <- harmonize_ucla_deaths(agencies = 'MA')

# Show variables present in MA decedent data
colnames(ma.deaths)

# UCLA.ID is present - join UCLA facility data
ma.deaths %>%
    pull_ucla_fac_data()

# Compare UCLA data with BJS data
compare_ucla_bjs() %>%
    subset(State == 'Massachusetts') %>%
    arrange(desc(Year))

# Plot MA decedent data (wrangled to monthly numbers)
ggplot() +
    geom_line(data = ma.deaths %>% group_by(Year, Month) %>% summarise(Total.Deaths = n()) %>% 
                  mutate(Date = as.Date(str_c(Year,Month,'01'), format = '%Y%B%d')),
              aes(x = Date, y = Total.Deaths)
              ) +
    geom_smooth(data = ma.deaths %>% group_by(Year, Month) %>% summarise(Total.Deaths = n()) %>% 
                  mutate(Date = as.Date(str_c(Year,Month,'01'), format = '%Y%B%d')),
              aes(x = Date, y = Total.Deaths)
              )

```


## Demographic Data

UCLA obtains custodial demographic data for Massachusetts state prisons from annual population reports from the Massachusetts Department of Correction.

Example files in corresponding folder and links to reports in raw data file.

Some processing of the files takes place. 

```{r}
# Load MA demographic data
ma.demographics <- harmonize_ucla_dem(agencies = 'MA')

# Show MA demographic data to assess origin type
head(ma.demographics)

# The origin type is combined. This means that age and sex groups were aggregated and reported together.  

# Plot MA demographic data 
ggplot() +
    geom_bar(data = ma.demographics %>% group_by(Date, Age) %>% summarise(Number = sum(Number, na.rm = TRUE)),
             aes(x = Date, y = Number, fill = Age), stat = 'identity') +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) # To see dates on x axis



```
