---
title: "WV_source_guide"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Environment

```{r}
# Load relevent packages and functions
library(tidyverse)
source('Code/general_utilities.R') # Note you may have to change your Rmarkdown option for evaluate chunks to project instead of document
options(scipen=10)
```

## Decedent Data

UCLA obtains custodial decedent data for West Virginia state prisons and jails (WV is a unified carceral system) from public records requests to the West Virginia Department of Corrections.

Example files included in corresponding documents folder. 

Some processing of the files takes place, including the selection of corresponding variables and linking data to UCLA datasets. 

```{r}
# Load WV decedent data
wv.deaths <- read_CMP_deaths(agencies = 'WV')

# Show variables present in WV decedent data
colnames(wv.deaths)

# UCLA.ID is present - join UCLA facility data
wv.deaths %>%
    pull_CMP_fac_data()

# Compare UCLA data with BJS data
compare_CMP_bjs(source = 'MCI') %>%
    subset(State == 'West Virginia') %>%
    arrange(desc(Year))

# Plot WV decedent data 
ggplot() +
    geom_line(data = wv.deaths %>% group_by(Year) %>% summarise(Total.Deaths = n()),aes(x = Year, y = Total.Deaths))

```


## Demographic Data

UCLA currently does not have demographic data processed for West Virginia state prisons and jails. Please wait for the next update to this data. 