---
title: "KY_source_guide"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Environment

```{r}
# Load relevent packages and functions
library(tidyverse)
source('Code/general_utilities.R') # Note you may have to change your Rmarkdown option for evaluate chunks to project instead of document
options(scipen=10)
```

## Decedent Data

UCLA obtains custodial decedent data for Kentucky state prisons from public records requests to the Kentucky Department of Corrections.

Example files included in corresponding documents folder. 

Some processing of the files takes place, including the selection of corresponding variables and linking data to UCLA datasets. 

```{r}
# Load KY decedent data
ky.deaths <- read_CMP_deaths(agencies = 'KY')

# Show variables present in KY decedent data
colnames(ky.deaths)

# UCLA.ID is present - join UCLA facility data
ky.deaths %>%
    pull_CMP_fac_data()

# Compare CMP data with BJS data
compare_CMP_bjs(source = 'MCI') %>%
    subset(State == 'Kentucky') %>%
    arrange(desc(Year))

# Plot KY decedent data 
ggplot() +
    geom_line(data = ky.deaths %>% group_by(Year) %>% summarise(Total.Deaths = n()),aes(x = Year, y = Total.Deaths))

```


## Demographic Data

UCLA currently does not have demographic data processed for Kentucky state prisons. Please wait for the next update to this data. 
