---
title: "WY_source_guide"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Decedent Data

UCLA obtains custodial decedent data for Wyoming state prisons from public records requests to the Wyoming Department of Corrections.

Example files included in corresponding documents folder.

Some processing of the files takes place, including the selection of corresponding variables and linking data to UCLA datasets. 

```{r}
# Load WY decedent data
wy.deaths <- read_ucla_deaths(agencies = 'WY')

# Show variables present in WY decedent data
colnames(wy.deaths)

# No UCLA.ID is present 

# Compare UCLA data with BJS data
compare_ucla_bjs(source = 'MCI') %>%
    subset(State == 'Wyoming') %>%
    arrange(desc(Year))

# Plot SD decedent data 
ggplot() +
    geom_line(data = ak.deaths %>% group_by(Year) %>% summarise(Total.Deaths = n()),aes(x = Year, y = Total.Deaths))

```


## Demographic Data

UCLA currently does not have demographic data processed for Wyoming state prisons. Please wait for the next update to this data. 
