---
title: "SD_source_guide"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Decedent Data

UCLA obtains custodial decedent data for South Dakota state prisons from public records requests to the South Dakota Department of Corrections.

Example files included in corresponding documents folder.

Some processing of the files takes place, including the selection of corresponding variables and linking data to UCLA datasets. 

```{r}
# Load SD decedent data
sd.deaths <- read_CMP_deaths(agencies = 'SD')

# Show variables present in SD decedent data
colnames(sd.deaths)

# UCLA.ID is present - join UCLA facility data
sd.facility <- sd.deaths %>%
                pull_CMP_fac_data()
head(sd.facility)

# Compare CMP data with BJS data
compare_CMP_bjs(source = 'MCI') %>%
    subset(State == 'South Dakota') %>%
    arrange(desc(Year))

# Plot SD decedent data 
ggplot() +
    geom_line(data = sd.deaths %>% group_by(Year) %>% summarise(Total.Deaths = n()),aes(x = Year, y = Total.Deaths))

```


## Demographic Data

UCLA currently does not have demographic data processed for South Dakota state prisons. Please wait for the next update to this data. 
