---
title: "MO_source_guide"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Environment

```{r}
# Load relevent packages and functions
library(tidyverse)
source('Code/general_utilities.R') # Note you may have to change your Rmarkdown option for evaluate chunks to project instead of document
options(scipen=10)
```

## Decedent Data

UCLA obtains custodial decedent data for Missouri state prisons from public records requests to the Missouri Department of Correction.

Example files included in corresponding documents folder.

Some processing of the files takes place, including the selection of corresponding variables and linking data to UCLA datasets. 

```{r}
# Load MO decedent data
mo.deaths <- read_ucla_deaths(agencies = 'MO')

# Show variables present in MO decedent data
colnames(mo.deaths)

# UCLA.ID is not present - please wait for data update to join UCLA facility data

# Compare UCLA data with BJS data
compare_ucla_bjs() %>%
    subset(State == 'Missouri') %>%
    arrange(desc(Year))

# Plot MO decedent data (wrangled to monthly numbers)
ggplot() +
    geom_line(data = mo.deaths %>% group_by(Year, Month) %>% summarise(Total.Deaths = n()) %>% 
                  mutate(Date = as.Date(str_c(Year,Month,'01'), format = '%Y%B%d')),
              aes(x = Date, y = Total.Deaths)
              ) +
    geom_smooth(data = mo.deaths %>% group_by(Year, Month) %>% summarise(Total.Deaths = n()) %>% 
                  mutate(Date = as.Date(str_c(Year,Month,'01'), format = '%Y%B%d')),
              aes(x = Date, y = Total.Deaths)
              )

```


## Demographic Data

UCLA obtains custodial demographic data for Missouri state prisons from annual population reports from the Missouri Department of Correction.

Example files in corresponding folder and links to reports in raw data file.

Some processing of the files takes place. 

```{r}
# Load MO demographic data
mo.demographics <- harmonize_ucla_dem(agencies = 'MO')

# Show MO demographic data to assess origin type
head(mo.demographics)

# The origin type is combined. This means that age and sex groups were aggregated and reported together. 

# Plot MO demographic data 
ggplot() +
    geom_bar(data = mo.demographics %>% group_by(Date, Age) %>% summarise(Number = sum(Number, na.rm = TRUE)),
             aes(x = Date, y = Number, fill = Age), stat = 'identity') +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) # To see dates on x axis



```
