---
title: "GA_source_guide"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Environment

```{r}
# Load relevent packages and functions
library(tidyverse)
source('Code/general_utilities.R') # Note you may have to change your Rmarkdown option for evaluate chunks to project instead of document
options(scipen=10)
```

## Decedent Data

UCLA obtains custodial decedent data for Georgia state prisons from public records requests to the Georgia Department of Corrections.

Example files included in corresponding documents folder.

Some processing of the files takes place, including the selection of corresponding variables and linking data 
to UCLA datasets. 

```{r}
# Load GA decedent data
ga.deaths <- harmonize_ucla_deaths(agencies = 'GA')

# Show variables present in GA decedent data
colnames(ga.deaths)

# UCLA.ID is not present - please wait for data update to join UCLA facility data

# Compare UCLA data with BJS data
compare_ucla_bjs() %>%
    subset(State == 'Georgia') %>%
    arrange(desc(Year))

# Plot GA decedent data (wrangled to monthly numbers)
ggplot() +
    geom_line(data = ga.deaths %>% group_by(Year, Month) %>% summarise(Total.Deaths = n()) %>% 
                  mutate(Date = as.Date(str_c(Year,Month,'01'), format = '%Y%B%d')),
              aes(x = Date, y = Total.Deaths)
              )

```


## Demographic Data

UCLA obtains custodial demographic data for Georgia state prisons from monthly population reports from the Georgia Department of Corrections.

Example files in corresponding folder and links to reports in raw data file.

Some processing of the files takes place. 

```{r}
# Load GA demographic data
ga.demographics <- harmonize_ucla_dem(agencies = 'GA')

# Show GA demographic data to assess origin type
head(ga.demographics)

# The origin type is combined. This means that age and sex groups were aggregated and reported together.  

# Plot GA demographic data 
ggplot() +
    geom_bar(data = ga.demographics %>% group_by(Date, Age) %>% summarise(Number = sum(Number, na.rm = TRUE)),
             aes(x = Date, y = Number, fill = Age), stat = 'identity') +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) # To see dates on x axis



```