---
title: "LA_source_guide"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Environment

```{r}
# Load relevant packages and functions
library(tidyverse)
source('Code/general_utilities.R') # Note you may have to change your Rmarkdown option for evaluate chunks to project instead of document
options(scipen=10)
```

## Decedent Data

UCLA obtains custodial decedent data for Louisiana state prisons from Incarceration Transparency. All credit for data production and maintenance goes to Incarceration Transparency and any use of the Louisiana decedent data contained in this dataset should follow their and our data license agreements.

Link to Incarceration Transparency custodial death dataset: https://www.incarcerationtransparency.org/?page_id=11

Some processing of the files takes place, including the selection of corresponding variables and linking data 
to UCLA datasets. 

```{r}
# Load LA decedent data
la.deaths <- read_CMP_deaths(all.agencies = FALSE, agencies = 'LA')

# Show variables present in LA decedent data
colnames(la.deaths)

# UCLA.ID is not present - please wait for data update to join UCLA facility data

# Compare CMP data with BJS data
compare_CMP_bjs(source = 'MCI') %>%
    subset(State == 'Louisiana') %>%
    arrange(desc(Year))

# Plot LA decedent data 
ggplot() +
    geom_line(data = la.deaths %>% group_by(Year) %>% summarise(Total.Deaths = n()),aes(x = Year, y = Total.Deaths))

```


## Demographic Data

UCLA currently does not have demographic data processed for Louisiana state prisons. Please wait for the next update to this data. 

