---
title: "MI_source_guide"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Environment

```{r}
# Load relevent packages and functions
library(tidyverse)
source('Code/general_utilities.R') # Note you may have to change your Rmarkdown option for evaluate chunks to project instead of document
options(scipen=10)
```

## Decedent Data

UCLA obtains custodial decedent data for Michigan state prisons from public records requests to the Michigan Department of Correction.

Example files included in corresponding documents folder.

Some processing of the files takes place, including the selection of corresponding variables and linking data to UCLA datasets. 

```{r}
# Load MI decedent data
mi.deaths <- read_CMP_deaths(agencies = 'MI')

# Show variables present in MI decedent data
colnames(mi.deaths)

# UCLA.ID is present - join UCLA facility data
mi.deaths %>%
    pull_CMP_fac_data()

# Compare CMP data with BJS data
compare_CMP_bjs(source = 'MCI') %>%
    subset(State == 'Michigan') %>%
    arrange(desc(Year))

# Plot MI decedent data 
ggplot() +
    geom_line(data = mi.deaths %>% group_by(Year) %>% summarise(Total.Deaths = n()),aes(x = Year, y = Total.Deaths))

```


## Demographic Data

UCLA obtains custodial demographic data for Michigan state prisons from annual population reports from the Michigan Department of Correction.

Example files in corresponding folder and links to reports in raw data file.

Some processing of the files takes place. 

```{r}
# Load MI demographic data
mi.demographics <- harmonize_CMP_dem(agencies = 'MI')

# Show MI demographic data to assess origin type
head(mi.demographics)

# The origin type is distinct. This means that age groups were reported separately from sex groups for demographic data. In order to construct age and sex groups for a date from this distinct data, we applied reported sex ratios to each age group. This assumption is likely incorrect. 

# Plot MI demographic data 
ggplot() +
    geom_bar(data = mi.demographics %>% group_by(Date, Age) %>% summarise(Number = sum(Number, na.rm = TRUE)),
             aes(x = Date, y = Number, fill = Age), stat = 'identity') +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) # To see dates on x axis



```