---
title: "HI_source_guide"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Environment

```{r}
# Load relevent packages and functions
library(tidyverse)
source('Code/general_utilities.R') # Note you may have to change your Rmarkdown option for evaluate chunks to project instead of document
options(scipen=10)
```

## Decedent Data

UCLA obtains custodial decedent data for Hawaii state prisons from public records requests to the Hawaii Department of Public Safety.

Example files included in corresponding documents folder. 

Some processing of the files takes place, including the selection of corresponding variables and linking data to UCLA datasets. 

```{r}
# Load HI decedent data
hi.deaths <- read_CMP_deaths(agencies = 'HI')

# Show variables present in HI decedent data
colnames(hi.deaths)

# UCLA.ID is not present - please wait for data update to join UCLA facility data

# Compare CMP data with BJS data
compare_CMP_bjs(source = "MCI") %>%
    subset(State == 'Hawaii') %>%
    arrange(desc(Year))

# Plot HI decedent data 
ggplot() +
    geom_line(data = hi.deaths,aes(x = Year, y = Total.Deaths)) 

```


## Demographic Data

UCLA currently does not have demographic data processed for Hawaii state prisons. Please wait for the next update to this data. 
